---
title: "Exploratory Data Analysis"
author: "Amber Lin"
date: "`r Sys.Date()`"
format: 
  revealjs:
    theme: night
    transition: fade  # Smooth transition effect
    slide-number: true 
bibliography: references.bib
nocite: "@*"
---

# Exploratory Data Analysis of Three World Development Indcators
-- focusing on GDP, Life Expectancy, Unemployemtn Rate

## Introduction
- This presentation explores three key economic indicators:
  - **GDP per Capita**
  - **Life Expectancy**
  - **Unemployment Rate**

## Data Overview
- The dataset contains key economic indicators.
- First few rows:

```{r, echo=FALSE}
library(tidyverse)

# Load the dataset
df <- read.csv("wdi.csv")

# Show first few rows
library(knitr)
kable(head(df, 5), caption = "Sample of the Dataset")
```

## Summary Statistics
- Summary of GDP per Capita, Life Expectancy, and Unemployment Rate:

```{r, echo=FALSE}
library(knitr)
selected_columns <- c("gdp_per_capita", "life_expectancy", "unemployment_rate")
df_selected <- df %>% select(all_of(selected_columns))
summary_statistics <- summary(df_selected)
kable(summary_statistics, caption = "Summary Statistics of Key Indicators: GDP, Life Expectancy, and Unemployment Rate")
```

## GDP per Capita Distribution
- The data is **left-skewed**, meaning most countries have **GDP per capita below $50,000**.

```{r, echo=FALSE}
ggplot(df, aes(x = gdp_per_capita)) +
  geom_histogram(fill = "blue", bins = 30) +
  labs(title = "GDP per Capita Distribution", x = "GDP per Capita ($)", y = "Frequency")
```

## GDP vs Life Expectancy
- A **positive correlation** exists between **GDP per capita** and **life expectancy**.
- This suggests that **economic growth leads to longer life expectancy**.

```{r, echo=FALSE}
ggplot(df, aes(x = gdp_per_capita, y = life_expectancy)) +
  geom_point(alpha = 0.5) +
  labs(title = "Life Expectancy vs GDP per Capita", x = "GDP per Capita ($)", y = "Life Expectancy (Years)")
```


## Unemployment Rate & Education Expenditure
- There is a **slightly positive correlation** between **education expenditure and unemployment rate**.

```{r, echo=FALSE}
ggplot(df, aes(x = unemployment_rate, y = education_expenditure_gdp_share)) +
  geom_point(alpha = 0.5, color = "blue") +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(title = "Unemployment Rate vs Education Expenditure",
       x = "Unemployment Rate (%)",
       y = "Education Expenditure (% of GDP)") +
  theme_minimal()
```


## Conclusion
- **GDP per Capita is left-skewed**: Most countries have lower GDP.
- **Higher GDP → Higher Life Expectancy**: Economic development improves health.
- **Education Spending & Unemployment**: No strong relationship found.

## References
